---
title: "Reproducible Research - Project 1"
author: "Guynemer Cétoute"
date: "`r Sys.Date()`"
output: html_document
---

This doocument is to show the acquired skills from the course Reproducible Research
from the R Data Science hosted by Johns Hopkins University from Coursera.

## The steps to load the Data

```{r loading, echo = TRUE}
library(tidyverse)
activity <- read_csv("Downloads/repdata_data_activity/activity.csv")
View(activity)
```

## Question 1: Calculate the total number of steps taken per day

Before showing the total number of steps per day, let us create a new column to assign the name of the day.
```{r days, echo = TRUE}
activity$days <- wday(activity$date, label = T)
```

```{r NumberPerDay, echo = TRUE}
activity %>% filter(!is.na(steps)) %>% group_by(days) %>% summarise(Total = sum(steps))
```

## Question 2: Histogram of the total number of steps taken each day

```{r histogram_1, echo = TRUE}
ggplot(activity %>% filter(!is.na(steps)), aes(x = steps)) +
    geom_histogram(binwidth = 100, fill = "skyblue", color = "black", alpha = 0.7) +
    facet_wrap(~ days) + labs(title = "Histogram of Steps per Day by Day of the Week",
         x = "Number of Steps", y = "Frequency") + theme_minimal()
```
## Question 3: Mean and median number of steps taken each day

### To show the number of NA in the column steps
```{r NA_count, echo = TRUE}
activity %>% filter(is.na(steps)) %>% nrow()
```

### The Means number of steps per day

```{r mean_1, echo = TRUE}
activity %>% filter(!is.na(steps)) %>% group_by(days) %>% summarise(Average = mean(steps))

```

### The Median number of steps per day

```{r median_1, echo = TRUE}
activity %>% filter(!is.na(steps)) %>% group_by(days) %>% summarise(Median = median(steps))
```

## Question 4: Time series plot of the average number of steps taken

```{r plot_time_series, echo = TRUE}
activity %>% filter(!is.na(steps)) %>% group_by(date) %>% 
    summarise(avg_steps = mean(steps)) %>%
    ggplot(aes(x = date, y = avg_steps)) + geom_line(color = "blue") +
    labs(title = "Average Steps per Day", x = "Date", y = "Average Steps") +
    theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Question 5: Code to describe and show a strategy for imputing missing data

Here, I impute the missing data in the steps column by the mean of the column.

```{r impute, echo = TRUE}
activity_2 <- activity %>% mutate(steps = ifelse(is.na(steps),
                                                 mean(steps, na.rm = T), steps))
```

## Question 6: Histogram of the total number of steps taken each day after missing values are imputed

```{r histogram_2, echo = TRUE}
ggplot(activity_2, aes(x = steps)) +
     geom_histogram(binwidth = 100, fill = "skyblue", color = "black", alpha = 0.7) +
     facet_wrap(~ days) +
     labs(title = "Histogram of Steps per Day by Day of the Week \nAfter imputing the missing value in the steps column", x = "Number of Steps", y = "Frequency") +
    theme_minimal()
```

## Question 7: Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

Before plotting the the graph comparing weekdays with weekends, I have to create another column to separate the weekdays from the weekends.

```{r transform, echo = TRUE}
activity_2$daysType <- case_when(activity_2$days %in%
                                     c('Mon', 'Tue', 'Wed', 'Thu', 'Fri') ~ 'weekday',
                                 activity_2$days %in% c('Sat', 'Sun') ~ 'weekend')
```

Now, this is the plot.

```{r final_plot, echo = TRUE}
activity_summary <- activity_2 %>%
     group_by(interval, daysType) %>%
     summarise(avg_steps = mean(steps, na.rm = TRUE))

ggplot(activity_summary, aes(x = interval, y = avg_steps)) +
     geom_line() +
     facet_wrap(~ daysType, ncol = 1) +
     labs(title = "Average Steps per 5-Minute Interval (Weekdays vs Weekends)",
                       x = "5-Minute Interval", y = "Average Number of Steps") +
     theme_minimal() +
     theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

